Efficiently stable presentations from error-correcting codes

### 2025-10-07 daj-revision2/referee-response.txt

Yes this was a typo in the last line of the proof, we corrected it.

17. page 28, line 3: what you write is not true for a fixed u, but it is true on
average over u.

A: Corrected

18. page 29, line -5– -1: the conventions for the notation poly here (given
in footnote 7) do not match those of [JNV+22]. For example, on the
last line of page 23 of [JNV+22], you have an expression of the form
poly(ε, n−1) + e−Ω(r/m2) that becomes, in the conclusion of [JNV+22,
Theorem 4.1] poly(ε, n−1, e−Ω(r/m2)). And similarly all over [JNV+22].
So, the way you prove Theorem 4.1 has two problems: (a) you do not
prove that the presentation given in (22) is stable in the sense of Definition
1.2 (because you do not prove limε δ(ε) = 0) and (2) your convention of
poly contradicts your use of it.
When you write Yes, we are sure that the modulus goes to 0 with epsilon,
I agree that this is true, but where is it proven? I did not find the proof
written, neither in [JNV+22] nor in your submission. I thought that what
is proven in [JNV+22] is with δ(ε) ≤ C(mC + dC )(ε1/C + q−1/C ) for some
universal C, which is C(mC + dC )q−1/C > 0 for ε = 0.
If you want to stick to the convention that limε δ(ε) = 0, one cheap way is
to exploit Lemma 2.6 and [GH17] to explain that the modulus of stabil-
ity in Theorem 4.1 is something like min(C1ε, C2poly(ε, 1/q)) for a good
(=polynomial in the parameters) constant C2 and a bad (=exponential in
the parameters) constant C1.

A: You're right, we were overly sloppy there. Yes, your interpretation of the poly() notation is correct and we made a mistake in translating it. This has been corrected. We added discussion around the fact that the bound does not go to 0 with epsilon.

[T: I am not super happy with this. The discussion is the paragraph right after the statement of Theorem 4.1. Does someone want to look at it and see if they manage to make it better?}

19. page 29, line -2: in [JNV+22], there is an assumption r ≥ 12mt (which
in the notation here becomes r ≥ 12m(d + 1)). Do you have assumptions
that ensure that Ω(tm2) satisfies this inequality?

A: If the constant implicit in the Omega(.) is at least 12, the condition is satisfied? We thought this should be clear.

20. page 29, line -1: you could explain why the existence of Gc implies the
conclusion of the theorem: write the homomorphism that is δ(ε)-close to

A: We added it.

21. Here is what I had in mind in my comment 35.

[Michael, I think that this is for you]

### 2025-10-07 daj-revision2/efficient_stability.tex
finished comments

\begin{theorem}\label{thm:z2-stab}
Let $\mC$ be the class of all tracial von Neumann algebras. 
The presentation of  $\Z_2^k$ given in~\eqref{eq:z2k-eff}, together with the distributions $\mu_S$ and $\mu_R$ defined above, is $(\delta,\mC)$ stable with \tnote{edited, referee did not like $\poly$}$\delta(\eps)= C (mdt)^{c_1} (\eps/q)^{c_2}$, where $C,c_1,c_2$ are universal positive constants.  
\end{theorem}

We briefly explain a possible setting of parameters in Theorem~\ref{thm:z2-stab}. For any integer $t\geq 1$, fix $q=2^t$ and let $d=m=c t^c$ for some constant $c>0$. Then $k=t(d+1)^m = 2^{\Theta((\log q)^c \log\log q)}$ and $N=q^{m+1}= 2^{\Theta((\log q)^{c+1} \log\log q)}$. Moreover, the number of relations in~\eqref{eq:z2k-eff} is $O(N^2)$, which scales as $2^{\poly\log k}$; and the maximum length of a relation is $d+2=O(\log k)$. Finally, with this choice of parameters the function $\delta(\eps)$ scales as $\poly(\log k)\cdot \poly(\eps)$. We have thus obtained a presentation of $\Z_2^k$ of size only mildly superpolynomial in $k$, and with soundness that depends polylogarithmically on $k$.

@@

We are now in a position to apply~\cite[Theorem 4.1]{ji2022quantum}. For this we need to define a synchronous strategy in the tensor code test $\code^{\otimes m}$, where $\code$ is the Reed-Solomon code with degree $d$ over $\F_q$, and thus $\code^{\otimes m}$ is the code $\code_\RM$ considered in Section~\ref{sec:rmq} (see the remark right after Lemma~\ref{lem:schwartz-zippel}). For the ``points measurement'' $A^u$ we choose $P^u$. For the ``lines measurement'' $B^\ell$ we choose $Q^\ell$ from Claim~\ref{claim:z2-stab-4}. Finally, for the ``pair measurement'' $P^{u,v}$ we choose $R^{u,v}$ from Claim~\ref{claim:z2-stab-3}. By Claim~\ref{claim:z2-stab-4} this strategy succeeds with probability $1-\poly(d,\eps)$ in the ``axis-parallel lines test'', and by Claim~\ref{claim:z2-stab-3} it succeeds with probability $1-\poly(\eps)$ in the ``subcube commutation test.'' Applying~\cite[Theorem 4.1]{ji2022quantum} we deduce the existence of a projective measurement $\{G_c\}_{c\in\code^{\otimes m}}$ on $\mA$ such that for all integers $r \geq 12mt$,
\[ \Es{u\in\F_q^m} \sum_{c\in\code^{\otimes m}} \tau\big( G_c P^u_{c(u)}\big) \geq 1-\eta\;,\]
where $\eta = \poly(m,d,r) \cdot\poly(\eps,q^{-1},e^{-\Omega(r/m^2)})$. Choosing $r=\Omega(m^2 t)$, since $q=2^{t}$, the bound becomes $\poly(m,d,t) \cdot\poly(\eps,q^{-1})$.

>>>>

\begin{theorem}\label{thm:z2-stab}
Let $\mC$ be the class of all tracial von Neumann algebras. 
The presentation of  $\Z_2^k$ given in~\eqref{eq:z2k-eff}, together with the distributions $\mu_S$ and $\mu_R$ defined above, is $(\delta,\mC)$ stable with $\delta(\eps)= C (mdt)^{c_1} \eps^{c_2}(1/q)^{c_3}$, where $C,c_1,c_2,c_3$ are universal positive constants.  
\end{theorem}

We briefly explain a possible setting of parameters in Theorem~\ref{thm:z2-stab}. For any integer $t\geq 1$, fix $q=2^t$ and let $d=m=c t^c$ for some constant $c>0$. Then $k=t(d+1)^m = 2^{\Theta((\log q)^c \log\log q)}$ and $N=q^{m+1}= 2^{\Theta((\log q)^{c+1} \log\log q)}$. Moreover, the number of relations in~\eqref{eq:z2k-eff} is $O(N^2)$, which scales as $2^{\poly\log k}$; and the maximum length of a relation is $d+2=O(\log k)$. Finally, with this choice of parameters the function $\delta(\eps)$ scales as $\poly(\log k)\cdot (\poly(\eps) + \poly(1/k))$. \tnote{added} While this bound is only good when $\eps$ is much larger than $1/k$,\footnote{If one looks closely the $\poly(1/k)$ term is in fact $2^{-\poly\log k}$.} and in particular it does not necessarily go to $0$ as $\eps\to 0$, this can be remedied by  
combining the bound we have obtained with the general Theorem~\ref{thm:gh} (which has a much larger prefactor). In this regime of not too small $\eps$, we thus obtain a presentation of $\Z_2^k$ of size only mildly superpolynomial in $k$, and with soundness that depends polylogarithmically on $k$. For smaller $\eps$ one may further adjust the parameters, e.g.\ take a larger $q$ than above but smaller $d$ and $m$ ; this will lead to alternate trade-offs (based on different presentations, since the presentation depends on $q$).

@@

We are now in a position to apply~\cite[Theorem 4.1]{ji2022quantum}. For this we need to define a synchronous strategy in the tensor code test $\code^{\otimes m}$, where $\code$ is the Reed-Solomon code with degree $d$ over $\F_q$, and thus $\code^{\otimes m}$ is the code $\code_\RM$ considered in Section~\ref{sec:rmq} (see the remark right after Lemma~\ref{lem:schwartz-zippel}). For the ``points measurement'' $A^u$ we choose $P^u$. For the ``lines measurement'' $B^\ell$ we choose $Q^\ell$ from Claim~\ref{claim:z2-stab-4}. Finally, for the ``pair measurement'' $P^{u,v}$ we choose $R^{u,v}$ from Claim~\ref{claim:z2-stab-3}. By Claim~\ref{claim:z2-stab-4} this strategy succeeds with probability $1-\poly(d,\eps)$ in the ``axis-parallel lines test'', and by Claim~\ref{claim:z2-stab-3} it succeeds with probability $1-\poly(\eps)$ in the ``subcube commutation test.'' Applying~\cite[Theorem 4.1]{ji2022quantum} we deduce the existence of a projective measurement $\{G_c\}_{c\in\code^{\otimes m}}$ on $\mA$ such that for all integers $r \geq 12mt$,
\[ \Es{u\in\F_q^m} \sum_{c\in\code^{\otimes m}} \tau\big( G_c P^u_{c(u)}\big) \geq 1-\eta\;,\]
where $\eta = \poly(m,d,r) \cdot\poly(\eps,q^{-1},e^{-\Omega(r/m^2)})$. Choosing $r=\Omega(m^2 t)$, since $q=2^{t}$, the bound becomes $\poly(m,d,t) \cdot\poly(\eps,q^{-1})$. The theorem is then proved where explicitly, the homomorphism that is close to $\phi$ is given by
\[ g: s_{u,a} \mapsto \sum_{\beta\in\F_q} (-1)^{a\cdot \kappa(\beta)} \sum_{c:c(u)=\beta} G_c \;.\]
\end{proof}
